Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates

### 2025-06-05 performance_comparison.tex
[MGa] subsec 5.1

\begin{figure}
	\centering
	\includegraphics{Fig/track.pdf}
	\caption{Sector of the Catalunya circuit considered in the analysis, corresponding to the curvilinear abscissa interval $\left[0.70, 0.77\right]$. Checkpoints are indicated by labels and are uniformly spaced by $\De\al=0.01$. This sector includes two distinct corners: a low-speed turn from $\al = 0.72$ to $\al = 0.73$, and a high-speed turn from $\al = 0.75$ to $\al = 0.76$, allowing for the evaluation of vehicle behavior across different dynamic scenarios.}
	\label{fig:track}
\end{figure}

\subsection{Open-loop parameters sensitivity}
\label{sec:ol_param_sensitivity}
To gain deeper insight into how key parameters affect the planning outcome, we focus here on the open-loop formulation.
%In this section, we explore the open-loop approach to gain deeper insight into the influence of key parameters on the results.
At this stage, only the track limit constraint is robustified. 
%For this analysis, we consider optimizations in which only the track limit constraint is robustified. 
This choice is motivated by the fact that its effects are more visually evident than those of the friction limit constraint, thereby facilitating a more immediate understanding of their impact.

@@

Observing the left panel of Figure~\ref{fig:ol_sensitivities}, it is evident that the trajectories corresponding to configurations with lower values of $\ga^\textrm{TLC}$ tend to travel closer to the track boundaries. In contrast, the trajectory associated with $\ga^\textrm{TLC} = 3$---represented by the light green line---remains significantly farther from the edges.
This behavior is consistent with the increased conservativeness introduced by higher values of $\ga^\textrm{TLC}$, which amplify the back-off term $\be^\textrm{TLC}$ and thereby enforce a larger safety margin from the track limits.
Chasing a trajectory associated with a larger safety margin leads to higher sector times; spanning $\ga^\textrm{TLC}$ in the range $\left[0,3\right]$ results in sector times that vary from 12.135\,s to 13.162\,s.

\begin{figure}
	\centering
	\includegraphics{Fig/gamma_sensitivity.pdf}
	\hfill
	\includegraphics{Fig/Pzero_sensitivity.pdf}
	\caption{Effect of $\ga^\textrm{TLC}$ (left panel) and $P_0$ (right panel) on optimal trajectory. The parameter $\ga^\textrm{TLC}$ spans in the range $\left[0,3\right]$ corresponding to a probability to meet the track limit constraint ranging from 50\% to 99\%. In the right panel, in addition to the baseline value $\bar{\bP}_0$ (black line), two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ (blue line) and $\bP_0 = \frac{1}{4}\bar{\bP}_0$ (red line). These correspond, respectively, to doubling and halving the initial uncertainty associated with the state vector.}
	\label{fig:ol_sensitivities}
\end{figure}

The second parameter examined is the initial value of the covariance matrix, $\bP_0$. 
We recall that at each grid point a new prediction horizon is initialized, and the corresponding matrix $\bP$ at its starting point, i.e., $\bP^0_k$, is set to $\bP_0$. 
The diagonal elements of $\bP$ are the variances $\text{var}(x_i)$ of each state variables, while the off-diagonal elements encode their covariances $\text{cov}(x_i,x_j)$. Larger diagonal values indicate higher uncertainty in the corresponding states, and nonzero off-diagonal terms imply mutual dependence between them. For simplicity, a diagonal $\bP_0$ is used in this study, that is, for each prediction horizon we assign an initial standard deviation $\sigma_i = \sqrt{\text{var}(x_i)}$ to each state and assume all initial covariances to be zero.

To provide a concrete interpretation, consider the first state variable $x_1$ with mean $\mu_1$ and standard deviation $\sigma_1$.
Under the Gaussian assumption, the true value of $x_1$ lies within the interval $\left[\mu_1 - \sigma_1,\ \mu_1 + \sigma_1\right]$ with approximately 68\% probability, and within the interval $\left[\mu_1 - 3\sigma_1,\ \mu_1 + 3\sigma_1\right]$ with approximately 99\% probability.

We assign a baseline initial covariance matrix $\bP_0 = \bar{\bP}_0$ by selecting standard deviations $\bar{\sigma}_i$ for each state variable. These standard deviations have been chosen based on the typical range of variation of each state.
Assuming the state vector is ordered as described in Section~\ref{sec:vehicle_model}, the values used are $\bar{\boldsymbol{\sigma}}~=~\left[0.1\,\textrm{m/s}, 0.01\,\textrm{m/s}, 0.01\,\textrm{rad/s}, 1\,\textrm{m}, 1\,\textrm{m}, 0.0175\,\textrm{rad}\right]^T$.
In addition to the baseline value $\bar{\bP}_0$, two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ and $\bP_0 = \frac{1}{4}\bar{\bP}_0$. These correspond, respectively, to doubling and halving the initial standard deviations $\sigma_i$ associated with the state vector.

>>>>

\begin{figure}
	\centering
	\includegraphics{Fig/track.pdf}
	\caption{The analyzed Catalunya circuit sector spans $\alpha \in [0.70, 0.77]$, with labeled checkpoints every $\Delta\alpha = 0.01$. It contains two corners -- a low-speed turn ($\alpha = 0.72-0.73$) and a high-speed turn ($\alpha = 0.75-0.76$) -- providing contrasting dynamic conditions for vehicle behavior evaluation.}
	\label{fig:track}
\end{figure}

\subsection{Open-loop parameter sensitivity}
\label{sec:ol_param_sensitivity}
To gain deeper insight into how key parameters affect the planning outcome, we focus here on the open-loop formulation.
%In this section, we explore the open-loop approach to gain deeper insight into the influence of key parameters on the results.
At this stage, only the track limit constraint is robustified.
%For this analysis, we consider optimizations in which only the track limit constraint is robustified.
This choice is motivated by the fact that its effects are more visually evident than those of the friction limit constraint, thereby facilitating a more immediate understanding of their impact.

@@

Observing the left panel of Figure~\ref{fig:ol_sensitivities}, it is evident that the trajectories corresponding to configurations with lower values of $\ga^\textrm{TLC}$ tend to travel closer to the track boundaries. In contrast, the trajectory associated with $\ga^\textrm{TLC} = 3$\,--- represented by the light green line \,--- remains significantly farther from the edges.
This behavior is consistent with the increased conservativeness introduced by higher values of $\ga^\textrm{TLC}$, which amplify the back-off term $\be^\textrm{TLC}$ and thereby enforce a larger safety margin from the track limits.
Chasing a trajectory associated with a larger safety margin leads to higher sector times; spanning $\ga^\textrm{TLC}$ in the range $\left[0,3\right]$ results in sector times that vary from 12.135\,s to 13.162\,s.

\begin{figure}
	\centering
	\includegraphics{Fig/gamma_sensitivity.pdf}
	\hfill
	\includegraphics{Fig/Pzero_sensitivity.pdf}
\caption{Effect of the TLC penalty coefficient $\gamma^{\mathrm{TLC}}$ (left panel) and the initial covariance $\mathbf P_{0}$ (right panel) on the optimal trajectory.
Raising $\gamma^{\mathrm{TLC}}$ from 0 to 3 increases the probability of satisfying the track-limit constraints from 50 \% to 99 \%.
The right panel compares the nominal covariance $\bar{\mathbf P}_{0}$ (black) with $4\bar{\mathbf P}_{0}$ (blue line) and $\tfrac14\,\bar{\mathbf P}_{0}$ (red line), i.e., twice and half the nominal state uncertainty.
}
%\caption{Effect of $\ga^\textrm{TLC}$ (left panel) and $\bP_0$ (right panel) on optimal trajectory. The parameter $\ga^\textrm{TLC}$ spans in the range $\left[0,3\right]$ corresponding to a probability to meet the track limit constraint ranging from 50\% to 99\%. In the right panel, in addition to the baseline value $\bar{\bP}_0$ (black line), two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ (blue line) and $\bP_0 = \frac{1}{4}\bar{\bP}_0$ (red line). These correspond, respectively, to doubling and halving the initial uncertainty associated with the state vector.}
	\label{fig:ol_sensitivities}
\end{figure}

The second parameter examined is the initial value of the covariance matrix $\bP_0$.
We recall that at each grid point a new prediction horizon is initialized, and the corresponding matrix $\bP$ at its starting point, i.e., $\bP^0_k$, is set to $\bP_0$.
The diagonal elements of $\bP$ are the variances $\Var{x_i}$  of each state variables, while the off-diagonal elements encode their covariances $\Cov{x_i}{x_j}$. Larger diagonal values indicate higher uncertainty in the corresponding states, and nonzero off-diagonal terms imply mutual dependence between them. For simplicity, a diagonal $\bP_0$ is used in this study, that is, for each prediction horizon we assign an initial standard deviation $\sig_{x_i}$ to each state and assume all initial covariances $\Cov{x_i}{x_j}$  to be zero.

 Assuming the state vector is ordered as described in Section~\ref{sec:vehicle_model}, the baseline initial covariance matrix is set to $\bP_0=\diag(\bbsig^2)$, with
\begin{align}
\bar{\boldsymbol{\sigma}}~=~\left[0.1\,\textrm{m/s}, 0.01\,\textrm{m/s}, 0.01\,\textrm{rad/s}, 1\,\textrm{m}, 1\,\textrm{m}, 0.0175\,\textrm{rad}\right]^T.
\end{align}
These standard deviations have been chosen based on the typical range of variation of each state. To provide a concrete interpretation at the starting condition, the assumption on first state $u \sim \calN(\mu_u,\sig_u^2)$, with $\sig_u=0.1$ m/s, implies that
$\Pr\{\mu_u-\sig_u \leq u \leq \mu_u+\sig_u\}\approx 68\%$ and $\Pr\{\mu_u-3\sig_u \leq u \leq \mu_u+3\sig_u\}\approx 99\%$.

In addition to the baseline value $\bar{\bP}_0$, two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ and $\bP_0 = \frac{1}{4}\bar{\bP}_0$. These correspond, respectively, to doubling and halving the initial standard deviations $\sig_{x_i}$ associated with each state vector.
